\documentclass[11pt]{article}
\usepackage{minted}
\usepackage{amsfonts, amssymb, amsmath, float}
\usepackage{enumerate, esint, nicefrac, algorithm2e}
\parindent 0px
\date{November 30, 2023}
\title{CS301 :: Homework 6}
\author{Ryan Magdaleno}

% Helpful ::
% \line(1,0){358px}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Problem 1. A FINITE Language}

Let $FINITE_{TM} = \{\langle M\rangle \,|\, M$ is a TM and $M$ accepts 
a finite language$\}$. \\
Prove that $FINITE_{TM}$ is undecidable.

\vspace{5px}\textbf{Solution ::}

Suppose $FINITE_{TM}$ is deciable. \\
Therefore there is a TM $C$ that decides it.

\vspace{5px}$A = $ "On input $\langle M, w\rangle$,

\hspace{30px}construct $M'=$ "on input $x$,

\hspace{45px}Simulate $M$ on input $w$.

\hspace{45px}If $M$ halts, reject, else accept."

\hspace{30px}Pass $M'$ to $C$.

\hspace{30px}If $C$ accepts $M'$, reject, else accept."

\vspace{5px}$C$ decides similar to $A_{TM}$ which is undecidable, a contradiction. \\
$\therefore FINITE_{TM}$ is undecidable.
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Problem 2. There Can Only Be ONE}

Let $ONE_{TM} = \{\langle M\rangle \,|\, M$ is a TM that accepts exactly 1 string$\}$.
\\Prove that $ONE_{TM}$ is undecidable.

\vspace{5px}\textbf{Solution ::}

Suppose for the sake of contradiction that $ONE_{TM}$ is deciable. Therefore a TM
$S$ decides it.

\vspace{5px}$A = $ "On input $\langle M, w\rangle$,

\hspace{30px}Construct $M' =$ "On input $x$,

\hspace{45px}If $x\ne w$, reject,

\hspace{45px}else, return $M(w)$."

\hspace{30px}Pass $M'$ to $S$. If $S$ accepts $M'$, accept, else reject."

\vspace{5px}TM $A$ decides $A_{TM}$ which is undecidable, we have a
contradiction. \\
$\therefore$ $ONE_{TM}$ is undecidable.

\line(1,0){358px}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Problem 3. A LENGTH-y Decision}

Let $LENGTHK_{TM} = \\\{\langle M, k\rangle \,|\, M$ is a TM that accepts all
length-$k$ strings$\}$. \\
Prove that $LENGTHK_{TM}$ is undecidable.

\vspace{5px}\textbf{Solution ::}

Suppose for the sake of contradiction that $LENGTHK_{TM}$ is deciable.\\
Therefore a TM $Z$ decides it.

\vspace{5px}$A =$ "On input $\langle M, w\rangle$,

\hspace{30px}Construct $M' =$ "On input $k$,

\hspace{45px}If $w$ does not have length $k$, reject.

\hspace{45px}else, return $M(w)$."

\hspace{30px}Pass $\langle M', k\rangle$ to $Z$ and run this machine for
$LENGTHK_{TM}$, 

\hspace{30px}if $M'$ accepts, $Z$ accepts, else $Z$ rejects."

\vspace{5px}TM $Z$ decides similar to $A_{TM}$ which is undecidable, a contradiction. \\
$\therefore LENGTHK_{TM}$ is undecidable.

\end{document}