\documentclass[11pt]{article}
\usepackage{minted}
\usepackage{amsfonts, amssymb, amsmath, float}
\usepackage{enumerate, esint, nicefrac, algorithm2e}
\parindent 0px
\date{\today}
\title{CS301 :: Homework 4}
\author{Ryan Magdaleno}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Problem 1. Context Free Grammars}

Produce the CFG for the following language $L$. Let $\sum=\{a,b,c,d\}$
$$L = \left\{c^{w+x}a^{y+x}b^zc^zd^{w+x} : w, x\ge0;y,z\ge1\right\}$$
\vspace{5px}\textbf{Solution ::}
\begin{align}
    S&\longrightarrow cSd \,|\, A \\
    A&\longrightarrow caA \,|\, B \\
    B&\longrightarrow aBd \,|\, aCd \\
    C&\longrightarrow bCc \,|\, bc
\end{align}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Problem 2. Chomsky Normal Form} 

Convert the following grammar into CNF, $\sum = \{0,1\}.$ \\
Show your work at each of the five steps: \\
\textbf{START, BIN, DEL, UNIT, TERM.}

$S\longrightarrow 0XY \,|\, Y$ \\
$X\longrightarrow 1X \,|\, Y$ \\
$Y\longrightarrow S \,|\, 0 \,|\, \epsilon$

\vspace{5px}\textbf{Solution ::}

\textbf{START ::}
\begin{align}
    S_0&\longrightarrow S \\
    S&\longrightarrow 0XY \,|\, Y \\
    X&\longrightarrow 1X \,|\, Y \\
    Y&\longrightarrow S \,|\, 0 \,|\, \epsilon
\end{align}
\line(1,0){358px}

\textbf{BIN ::}
\begin{align}
    S_0&\longrightarrow S\\
    S&\longrightarrow 0S_1 \,|\, Y \\
    S_1&\longrightarrow XY\\
    X&\longrightarrow 1X \,|\, Y \\
    Y&\longrightarrow S \,|\, 0 \,|\, \epsilon
\end{align}
\line(1,0){358px}

\textbf{DEL ::}
\begin{align}
    S_0&\longrightarrow S \,|\, \epsilon \\
    S&\longrightarrow 0S_1 \,|\, Y \,|\, 0 \\
    S_1&\longrightarrow XY \,|\, X \,|\, Y \\
    X&\longrightarrow 1X \,|\, Y \,|\, 1 \\
    Y&\longrightarrow S \,|\, 0
\end{align}
\pagebreak

\textbf{UNIT ::}
\begin{align}
    S_0&\longrightarrow \epsilon \,|\, 0S_1 \,|\, 0 \\
    S&\longrightarrow 0S_1 \,|\, 0 \\
    S_1&\longrightarrow XY \,|\, 1X \,|\, 0S_1 \,|\, 1 \,|\, 0 \\
    X&\longrightarrow 1X \,|\, 0S_1 \,|\, 1 \,|\, 0 \\
    Y&\longrightarrow 0S_1 \,|\, 0
\end{align}
\line(1,0){358px}

\textbf{TERM ::}
\begin{align}
    S_0&\longrightarrow \epsilon \,|\, U_0S_1 \,|\, 0 \\
    S&\longrightarrow U_0S_1 \,|\, 0 \\
    S_1&\longrightarrow XY \,|\, U_1X \,|\, U_0S \,|\, 1 \,|\, 0 \\
    X&\longrightarrow U_1X \,|\, U_0S_1 \,|\, 1 \,|\, 0 \\
    Y&\longrightarrow U_0S_1 \,|\, 0 \\
    U_0&\longrightarrow 0 \\
    U_1&\longrightarrow 1
\end{align}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textbf{Problem 3. Non-context free Proof}

Prove that the following language is not context free. Let $\sum = \{a,b,c\}$
$$L = \{a^nb^nc^i : n < i < 2n\}$$

\vspace{5px}\textbf{Solution ::}

I assume $n=0$ isn't possible because $0 < 0+1 < 2(0)$ isn't true. \\
I assume $n=1$ isn't possible because $0 < 1+1 < 2(1)$ isn't true.

\vspace{7px}Suppose, for the sake of contradiction, that $L$ is context free. Then by
definition, there must be a CFG $G$ with pumping length $p$ that generates it. \\
Let $s = a^pb^pc^{p+1}$. \\
By the pumping lemma, $s$ can be partitioned into $u, v, x, y, z$ such that $|vxy| \le p$
and $|vy|\ge 1$. We proceed on cases for $v$ and $y$:

\vspace{7px}If $v$ and $y$ consist of the same characters:
\begin{itemize}
\item
If $v$ and $y$ consist of only $a$'s or only $b$'s then $vy$ consist of $k$ $a$'s or
$k$ $b$'s, for some $k\ge 1$. Because $|vxy|\le p$, $vxy$ is contained within the
$p$ $a$'s or $p$ $b$'s. Then $uxz$ is equal to either $a^{p-k}b^pc^{p+1}$ or 
$a^pb^pc^{p-k+1}$, none of which are within $L$ when $v/y$ is pumped down,
thus $uv^0xy^0z$ is not within $L$, which is a contradiction.

\item 
If $v$ and $y$ consist of only $c$'s then $vy$ must consist of $k$ $c$'s, for some 
$k\ge 1$. Because $|vxy|\le p$, $vxy$ is contained within the $p+1$ $c$'s. Therefore
$uxz$ is equal to $a^pb^pc^{p-k+1}$. From this we can see that $c^{p-k+1}$ will
make it not be contained within $L$ due to $k$ needing to be $\ge 1$ when we
pump down, this we've reached a contradiction.
\end{itemize}

\vspace{7px}If $v$ and $y$ consist of different characters:
\begin{itemize}
\item 
If $v$ consists of only $a$'s and $y$ consists of only $b$'s, for simplicity sake let's
say they contain the same number of their respective character, say $k$ characters
for each, for some $k\ge 1$. Because $|vxy|\le p$, $vxy$ should be contained within
the last $k$ characters of $a$, and the first $k$ characters of $b$.

From this we can see that $uxz$ is equal to $a^{p-k}b^{p-k}c^{p+1}$. From this now,
we can see that this will not be within $L$ if we pump down and are left with
$a^{p-k}b^{p-k}c^{p+1}$, this is not within $L$ because $k$ must be $\ge 1$,
causing a contradiction.

\item
If $v$ consists of only $b$'s and $y$ consists of only $c$'s, for simplicity sake let's
say they contain the same number of their respective characters, say $k$ characters
for each, for some $k\ge 1$. \\Because $|vxy|\le p$, $vxy$ should be contained within
the last $k$ \\characters of $b$, and the first $k$ characters of $c.$ From this
we can see that $uxz$ is equal to $a^pb^{p-k}c^{p-k+1}$, when we pump down we are
left with that result, which falls out of $L$, thus causing a contradiction.
\end{itemize}
All cases result in a contradiction of the pumping lemma.\\
$\therefore L$ is not context-free.

\end{document}